---
title: "Compositional Abstractions Analyses"
author: "Will McCarthy, Cameron Holdaway, Robert Hawkins"
date: "07/26/2023"
output: html_document
---

# Imports

## Packages

```{r setup, include=FALSE}
library(tidyverse)
library(lme4)
library(lmerTest)
library(tidyboot)
library(here)
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```

## Configuration

```{r ggplot, include=FALSE}
theme_set(theme_classic())
theme_update(# axis labels
  axis.title = element_text(size = 28),
  axis.text = element_text(size = 24),
  title = element_text(size = 24),
  legend.position = 'FALSE', 
  text = element_text(size=16), 
  element_line(size=1), 
  element_rect(size=2, color="#00000"))

primary_color = '#21606C'
secondary_color = '#22AAAA'

get_tower_set <- function(d) {
  d %>% 
    mutate(towerSet = case_when(
      (leftTarget %in% c('L', 'Pi')) & (rightTarget %in% c('L', 'Pi')) ~ 'L/Pi',
      (leftTarget %in% c('L', 'C')) & (rightTarget %in% c('L', 'C')) ~ 'L/C', 
      (leftTarget %in% c('Pi', 'C')) & (rightTarget %in% c('Pi', 'C')) ~ 'C/Pi'
    )) %>%
    return()
}

```

## Import data

These files are generated by `compabs_analysis_df_generator.ipynb`.
 
### Game-level exit survey data
 
```{r}
df_exit.raw = read_csv(here('results/csv/df_exit.csv'))
```

### Trial-level metadata

```{r}
df_trial.raw = read_csv(here('results/csv/df_trial.csv')) %>%
  get_tower_set() %>%
  mutate(trialNum = as.numeric(as.character(trialNum)),
         repNum = as.numeric(as.character(repNum)) + 1)
```

### Block placement data (builder side)

```{r}
df_block.raw = read_csv(here('results/csv/df_block.csv')) %>%
  get_tower_set() %>%
  mutate(repNum = as.numeric(as.character(repNum)),)
```

### Chat message data (architect side)

```{r}
df_chat.raw = read_csv(here('results/csv/df_chat.csv')) %>%
  mutate(trialNum = as.numeric(as.character(trialNum)),
         repNum = as.numeric(as.character(repNum)))
```

## Exclusion criteria 

Remove trials that don't reach criterion (75% Accuracy on 75% of trials).

It turns out that the exclusions of the first preregistered accuracy check is already a super set of the other exclusions.

```{r}
#How many dyads fulfill 
passedAcc <- df_trial.raw %>% 
  group_by(gameid, trialNum) %>% 
  filter(trialScore > 75  & repNum != 'practice') %>%
  group_by(gameid)%>% 
  tally %>% 
  filter(n/12 >=0.75) %>% 
  distinct(gameid)

# Note that 'yes' means 'yes I understood the instruction'
passedConfused <- df_exit.raw %>% 
  filter(confused == 'yes') %>% 
  distinct(gameid)

#How many dyads said they spoke English
passedEnglish <- df_exit %>% 
  filter(nativeEnglish == 'yes') %>% 
  distinct(gameid)


gamesToKeep = intersect(passedAcc, passedConfused) %>% intersect(passedEnglish) %>% pull(gameid)
```

### Count participants

```{r}
cat('before exclusions:', df_exit$gameid %>% unique() %>% length(), 'games started')
cat('after exclusions:', length(gamesToKeep), 'completed in dataset')
```

### Make final version of data frames

```{r}
df_trial <- df_trial.raw %>% 
  filter(!is.na(trialNum)) %>% 
  subset(gameid %in% gamesToKeep)

df_chat <- df_chat.raw %>% 
  filter(!is.na(trialNum)) %>% 
  subset(gameid %in% gamesToKeep)

df_block <- df_block.raw %>% 
  filter(!is.na(trialNum)) %>% 
  subset(gameid %in% gamesToKeep)

df_exit <- df_exit.raw %>% 
  subset(gameid %in% gamesToKeep)
```

# Results

```{r}
bootstrap_seed = 0
```

## Accuracy 

```{r}
set.seed(bootstrap_seed)

acc.indiv <- df_trial %>%
  group_by(repNum, gameid) %>%
  summarize(trialScore = mean(trialScore))

acc.boot <- acc.indiv %>%
  group_by(repNum) %>%
  tidyboot_mean(column=trialScore, nboot=1000)

acc.boot
```


```{r}
ggplot(acc.boot, aes(x=repNum, y=empirical_stat)) +
  geom_jitter(aes(y = trialScore), width = 0.1, height = 2, 
              data = acc.indiv, alpha = 0.2) +
  geom_line(size=1.5, color = primary_color) +
  geom_point(size = 4, color = primary_color) +
  geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), width = 0, size = 1.5,
               color = primary_color) +  
  labs(y = "accuracy (F1)", x = "repetition block", title = 'builder accuracy') +
  scale_y_continuous(limits = c(50,105), breaks = c(50,60,70,80,90,100),
                     labels=c('0.5','0.6','0.7','0.8','0.9','1.0')) +
  theme(legend.position = 'FALSE', 
        aspect.ratio = 1,
        text = element_text(size=16), element_line(size=1), 
        plot.title = element_text(hjust = 0.5, size=16), 
        element_rect(size=2, color="#00000"))
ggsave(here('results/plots/fig3A.pdf'), width=11, height = 11, units='cm', useDingbats = F)
```

### Mixed effects model

```{r}
m.acc <- df_trial %>%
  lmer(trialScore/100 ~ poly(repNum,2) + (1 + poly(repNum,2) | gameid),
       data = .)
summary(m.acc)
```


## Word Count 

```{r}
set.seed(bootstrap_seed)

wc.indiv <- df_trial %>%
  group_by(repNum, gameid) %>%
  summarize(word_count = mean(word_count)) 

wc.boot <- wc.indiv %>%
  group_by(repNum) %>%
  tidyboot_mean(column=word_count, nboot=1000)
wc.boot
```

### Figure 3B

```{r}
wc.boot %>%
  ggplot(aes(x=repNum,y=empirical_stat)) +
    geom_jitter(aes(y = word_count), data = wc.indiv, 
                alpha = 0.2,width =0.1)+
    geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), width = 0, size = 1.5,
                  data = wc.boot, color = primary_color) +  
    geom_line(size=1.5, color = primary_color)+
    geom_point(size = 4, color = primary_color) +
    labs( y = "words", x = "repetition", title = "words per trial") +
    theme(legend.position = 'FALSE', 
          plot.title = element_text(hjust = 0.5, size=16), 
          aspect.ratio = 1,
          text = element_text(size=16), 
          element_line(size=1),
          element_rect(size=2, color="#00000"))
ggsave(here('results/plots/fig3B.pdf'), width=11, height = 11, units='cm', useDingbats = F)
```

### Mixed effects model

```{r}
m.wc = df_trial %>% 
  lmer(word_count ~ poly(repNum, 2) + (1 + poly(repNum, 2) | gameid) + (1 + poly(repNum, 1)  | towerSet),
       data = .)
summary(m.wc)

# technically as an integer count variable >0, we should be using poisson
m.wc.poisson = df_trial %>% 
  glmer(word_count ~ poly(repNum, 2) + (1 + poly(repNum, 2) | gameid) + (0 + poly(repNum, 1) | towerSet),
       data = ., family = 'poisson')
summary(m.wc.poisson)
```


### Num messages 

#### Number of messages across trials
```{r}
set.seed(bootstrap_seed)

mess.boot <- df_trial %>%
  group_by(repNum) %>%
  tidyboot_mean(column=n_messages, nboot=1000)
mess.boot
```

```{r}
mess.boot %>%
  ggplot(aes(x=repNum, y=empirical_stat)) +
    geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), width = 0, size = 1.5,
                  data = mess.boot, color = primary_color) +  
    geom_line(size=1.5, color = primary_color)+
    geom_point(size = 4, color = primary_color) +
    ylab("messages") +
    xlab("repetition") +
    # theme_few() +
    # scale_x_continuous(breaks = round(seq(1, 4, by =1),1)) +
    xlim(c(0.75,4.25)) +
    # ylim(c(1,7)) +
    #scale_x_continuous(limits = c(1,6), breaks = c(1,2,3,4),labels=c('first', '2nd','3rd','final')) +
    scale_y_continuous(limits = c(1,6), breaks = c(1,2,3,4,5,6),labels=c(' 1', ' 2',' 3',' 4',' 5',' 6')) +
    ggtitle('messages per trial') +
    theme(legend.position = 'FALSE', text = element_text(size=16), element_line(size=1), plot.title = element_text(hjust = 0.5, size=16), element_rect(size=2, color="#00000"))
ggsave('./plots/unused/CA_ORLR20_messages_reps.pdf', width=7, height = 11, units='cm', useDingbats = F)
```

#### Linear mixed effects model for number of messages
```{r}
m.mes = lmer(data = df_trial, n_messages ~ repNum + (1 + repNum || gameid) + (1 | towerSet))
summary(m.mes)
```

### Timing 

Have to exclude participants marked as 'flagged', as techical errors/ connectivity issues mean they had some negative block placement times.

#### Total time to send messages across trials
```{r}
set.seed(bootstrap_seed)

chat.time.boot <- df_trial %>%
  group_by(repNum) %>%
  tidyboot_mean(column=total_chat_duration, nboot=100)
chat.time.boot
```

```{r}
chat.time.boot %>%
  ggplot(aes(x=repNum, y=empirical_stat)) +
    geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), width = 0, size = 1.5,color = primary_color) +  
    geom_line(size=1.5, color = primary_color)+
    geom_point(size = 4, color = primary_color) +
    ylab("total chat duration (s)") +
    xlab("repetition") +
    # theme_few() +
    # scale_x_continuous(breaks = round(seq(1, 4, by =1),1)) +
    # xlim(c(0.5,4.5)) +  
    # ylim(c(0,7)) +
    #scale_x_continuous(limits = c(0.6,4.4), breaks = c(1,2,3,4),labels=c('first', '2nd','3rd','final')) +
    scale_y_continuous(limits = c(40000,200000), labels = c('40','60','120','160','200')) +
    ggtitle('messaging time') +
    theme(legend.position = 'FALSE', text = element_text(size=16), element_line(size=1), plot.title = element_text(hjust = 0.5, size=16), element_rect(size=2, color="#00000"))
ggsave('./plots/CA_ORLR20_chat_duration_reps.pdf', width=7, height = 11, units='cm', useDingbats = F)
```

```{r}
m.chat.time = lmer(data = df_trial, total_chat_duration ~ repNum + (1 + repNum | gameid) + (1 | towerSet))
summary(m.chat.time)
```



#### Total time to send messages across trials
```{r}
set.seed(bootstrap_seed)

block.time.boot <- df_trial %>%
  group_by(repNum) %>%
  tidyboot_mean(column=total_block_duration, nboot=1000)
```

```{r}
block.time.boot %>%
  ggplot(aes(x=repNum, y=empirical_stat)) +
    geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), width = 0, size = 1.5, color = primary_color) +  
    geom_line(size=1.5, color = primary_color)+
    geom_point(size = 4, color = primary_color) +
    ylab("total build duration") +
    xlab("repetition") +
    theme_few() +
    theme(legend.position = 'FALSE', text = element_text(size=16), element_line(size=1), element_rect(size=2, color="#00000"))

```

```{r}
m.block.time = lmer(data = df_trial, total_block_duration ~ repNum + (1 + repNum | gameid) + (1 | towerSet))
summary(m.block.time)
```


## Referring expression analyses

```{r}
df_ref_exps = read_csv('../../results/csv/df_ref_exps.csv')
df_ref_exps_melt = read_csv('../../results/csv/df_ref_exps_melt.csv')
```


```{r}
m1 = lmer(data = df_ref_exps_melt, n_refs ~ exp_type + (1 | workerID) + (1 | tower_pair))
m2 = lmer(data = df_ref_exps_melt, n_refs ~ rep + exp_type + ( 1| workerID) + (1 | tower_pair))
m3 = lmer(data = df_ref_exps_melt, n_refs ~ rep*exp_type + (1 | workerID) + (1 | tower_pair))
# m4 = lmer(data = df_ref_exps_melt, n_refs ~ rep*exp_type + (1 | tower_pair))
# m5 = lmer(data = df_ref_exps_melt, n_refs ~ rep* exp_type + (1 | tower_pair))
anova(m1, m2, m3)
summary(m3)

```

```{r}
df_ref_exps_melt
```


## Read in Data
```{r}

#Sum ref exp then divide by 4 (since we have 4 raters)
dat_grouped = df_ref_exps_melt %>%  group_by(dyad_gameid, trial_num, rep, exp_type) %>% 
  summarise(sum_ref = sum(n_refs)/4) # 4 annotators

dat_grouped = spread(dat_grouped,exp_type	,sum_ref)

dat_grouped$mixType = ifelse(dat_grouped$block>0 & dat_grouped$tower > 0, 'mixed', ifelse(dat_grouped$block>0, 'block',ifelse(dat_grouped$tower>0,'tower','none')))
```

### Plot
```{r}

df_ref =  as.data.frame(table(dat_grouped$rep, dat_grouped$mixType))
df_ref
df_ref = df_ref %>% rename(repNum = Var1, mixType = Var2)

df_ref = df_ref %>% filter(mixType != 'none')

ggplot(df_ref, aes(x=repNum, y = Freq, fill=mixType)) + 
    geom_bar(stat="identity",position = 'fill') +
    scale_fill_manual(values=c('black', primary_color, "light grey")) +
  xlab("repetition") +
  ylab('frequency') +
  # ggtitle('builder accuracy') +
  theme(legend.position = 'FALSE', text = element_text(size=16), element_line(size=1), plot.title = element_text(hjust = 0.5, size=16), element_rect(size=2, color="#00000"))

ggsave("../../results/plots/referring_expressions_barplot.pdf", width=20, height = 26, units='cm', useDingbats = F)
```
```{r}
aggregate_trials <- df_ref_exps_melt %>% 
  group_by(dyad_gameid, trial_num, rep, exp_type, workerID) %>% 
  summarise(sum_refs = sum(n_refs))
aggregate_trials

aggregate_ratings <- aggregate_trials %>% 
  group_by(dyad_gameid, trial_num, rep, exp_type) %>% 
  summarise(mean_refs = mean(sum_refs))
aggregate_ratings

aggregate_reps <- aggregate_ratings %>% 
   group_by(rep, exp_type) %>% 
   summarise(mu = mean(mean_refs), sd = sd(mean_refs), n = n(), sem = sd(mean_refs)/sqrt(n()))
aggregate_reps
```



```{r}
ggplot(aggregate_reps, aes(x=rep, y=mu, colour=exp_type)) + geom_line(size = 1.5)+
  geom_errorbar(aes(ymin=mu-sem, ymax=mu+sem), width = 0, size = 1.5)+
  geom_point(size = 4) +
  # theme_few() +
  xlab("repetition")+
  ylab("referring expressions")+
  xlim(c(0.5,4.5)) +  
  theme(legend.position = 'FALSE', text = element_text(size=16),
        element_line(size=1),
        element_rect(size=2, color="#00000"))+
  scale_color_manual(values=c("#151515","#7D7D7D"))
# ggsave('./plots/CA_ORLR20_referring_exps.pdf', width=7, height = 11, units='cm', useDingbats = F)
# 
# 
# data(assemblage)
# chiperm(data=assemblage, B=199, resid=TRUE, cramer=TRUE)

```

### Proportion of trials that are perfect reconstructions

We might be worried about the stat above, since they're modeling a bounded F1 score with gaussian error.

We might instead be interested in a stricter accuracy metric, the binary variable corresponding to whether it was a perfect reconstruction or not.

```{r}
set.seed(bootstrap_seed)

perfect.boot <- df_trial %>%
  mutate(perfect = as.numeric(trialScore == 100)) %>%
  group_by(repNum) %>% 
  tidyboot_mean(column=perfect, nboot=1000)
```

```{r}
ggplot(perfect.boot, aes(x=repNum, y=empirical_stat)) +
  geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), 
                width = 0, size = 1.5, color=primary_color) +  
  geom_line(size=1.5, color=primary_color)+
  geom_point(size = 4, color=primary_color) +
  labs(y = "% perfect reconstruction", x = "Repetition block", title = 'stricter accuracy metric') +
  xlim(c(0.75,4.25)) +
  ylim(c(0.4,1)) +
  theme(legend.position = 'FALSE', text = element_text(size=16), element_line(size=1), 
        plot.title = element_text(hjust = 0.5, size=16), element_rect(size=2, color="#00000"))
```

```{r}
m.acc = df_trial %>% 
  mutate(perfect = trialScore == 100) %>%
  glmer(perfect ~ repNum + (1 + repNum | gameid),
        data = ., family = 'binomial')
summary(m.acc)
```

### Utterance length replicates with character count 

```{r}
set.seed(bootstrap_seed)

cc.boot <- df_trial %>%
  group_by(repNum) %>%
  tidyboot_mean(column=char_count, nboot=1000)
cc.boot
```


```{r}
cc.boot %>%
  ggplot(aes(x=repNum,y=empirical_stat)) +
    geom_errorbar(aes(ymin=ci_lower, ymax = ci_upper), width = 0, size = 1.5,
                  data = cc.boot, color = primary_color) +  
    geom_line(size=1.5, color = primary_color)+
    geom_point(size = 4, color = primary_color) +
    ylab("characters") +
    xlab("repetition") +
    # theme_few() +
    # scale_x_continuous(breaks = round(seq(1, 4, by =1),1)) +
    xlim(c(0.75,4.25)) +
    ylim(c(0,350)) +
    ggtitle('characters per trial') +
    #scale_x_continuous(limits = c(0.6,4.4), breaks = c(1,2,3,4),labels=c('first', '2nd','3rd','final')) +
    # scale_y_continuous(breaks = seq(-0.5,2.5,0.5)) +
    theme(legend.position = 'FALSE', text = element_text(size=16), element_line(size=1), plot.title = element_text(hjust = 0.5, size=16), element_rect(size=2, color="#00000"))
ggsave('./plots/unused/CA_ORLR20_char_count_reps.pdf', width=7, height = 11, units='cm', useDingbats = F)
```

```{r}
m.cc = lmer(data = df_trial, char_count ~ repNum + (1 + repNum || gameid) + (1 | towerSet))
summary(m.cc)
```

